\texttt{MSnbase}, efficient and elegant R-based processing and visualisation of raw mass spectrometry data

### 2020-07-22 msnbase2.Rnw
address comments ...

In \texttt{MSnbase}, mass spectrometry experiments are handled as
\texttt{MSnExp} objects. While the implementation is more complex, it
is useful to schematise a raw data experiment as being composed of raw
data, i.e. a collection of individual spectra, as well as
spectra-level metadata. Each spectrum is composed of m/z values and
associated intensities. The metadata are represented by a
\textcolor{red}{single} table with variables along the columns and
each row associated to a spectrum. Among the metadata available for
each spectrum, there are MS level, acquisition number, retention time,
precursor m/z and intensity (for MS level 2 and above), and many
more. \texttt{MSnbase} \textcolor{red}{relies on the \texttt{mzR}
  package~\cite{Chambers:2012} to import raw mass spectrometry data
  from one of the many community-maintained open standards formats
  (mzML, mzXML, mzData or ANDI-MS/netCDF) and} provides a rich
interface to manipulate such objects. The code chunk below illustrates
such an object as displayed in the R console and an enumeration of the
metadata fields.

@@

After processing of the BoxCar data, the final object can either be
further analysed in R or written back to disk as an \texttt{mzML} file
using \texttt{writeMSData()} for processing with other tools.

>>>>

In \texttt{MSnbase}, mass spectrometry experiments are handled as
\texttt{MSnExp} objects. While the implementation is more complex, it
is useful to schematise a raw data experiment as being composed of raw
data, i.e. a collection of individual spectra, as well as
spectra-level metadata (Figure \ref{fig:raw}). Each spectrum is
composed of m/z values and associated intensities. The metadata are
represented by a \textcolor{red}{single} table with variables along
the columns and each row associated to a spectrum. Among the metadata
available for each spectrum, there are MS level, acquisition number,
retention time, precursor m/z and intensity (for MS level 2 and
above), and many more. \texttt{MSnbase} \textcolor{red}{relies on the
  \texttt{mzR} package~\cite{Chambers:2012} to import raw mass
  spectrometry data from one of the many community-maintained open
  standards formats (mzML, mzXML, mzData or ANDI-MS/netCDF) and}
provides a rich interface to manipulate such objects. The code chunk
below illustrates such an object as displayed in the R console and an
enumeration of the metadata fields.

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{./figure/raw.png}
  \caption{Schematic representation of what is referred to by
    \textit{raw data}: a collection of mass spectra and a table
    containing spectrum-level annotations along the lines. Raw data is
    contained in one of the many community-maintained open standards
    formats (mzML, mzXML, mzData or ANDI-MS/netCDF). }
  \label{fig:raw}
\end{figure}

@@

After processing of the BoxCar data, the final object can either be
further analysed \sout{in R} \textcolor{red}{using \texttt{MSnbase}}
or written back to disk as an \texttt{mzML} file using
\texttt{writeMSData()} for processing with other tools.
